---
title: "Mixed Effects Models: Co-creation and Learning Outcomes"
author: "BuildBots Frame Paper Analysis"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dev = 'pdf'
)
```

# Overview

This analysis examines the relationship between co-creation level (during LLM interaction) and learning outcomes using mixed effects models. We account for the nested structure of students within groups.

**Research Questions:**

1. Does co-creation level predict post-test knowledge (T7)?
2. Does baseline knowledge (T1) predict co-creation level (selection effect)?
3. How does knowledge evolve over time (T1 → T4 → T7)?
4. Does the co-creation effect vary by timepoint?

# Load Libraries

```{r libraries}
library(tidyverse)      # Data manipulation and visualization
library(lme4)           # Mixed effects models
library(lmerTest)       # p-values for mixed models
library(sjPlot)         # Model visualization
library(performance)    # Model diagnostics
library(ggplot2)        # Advanced plotting
library(gridExtra)      # Multiple plots
library(knitr)          # Tables
library(broom.mixed)    # Tidy model outputs
```

# Load and Prepare Data

## Load Raw Data

```{r load-data}
# Load children knowledge data (T1, T4, T7)
df_children <- read_csv(
  '../Data/Qualtrics/BuildBots_children_T1T4T7.csv',
  show_col_types = FALSE
)

# Load engagement data to get Group assignments
df_engagement <- read_csv(
  '../Data/Qualtrics/BuildBots_engagement_children_T1-T7.csv',
  show_col_types = FALSE
)

# Load group-level co-creation data
df_cocreation <- read_csv(
  '../Data/BuildbotAnalysis - LLM analysis.csv',
  skip = 1,
  show_col_types = FALSE
)

# Display structure
cat("Children data dimensions:", dim(df_children), "\n")
cat("Engagement data dimensions:", dim(df_engagement), "\n")
cat("Co-creation data dimensions:", dim(df_cocreation), "\n")
```

## Extract Group Mapping and Merge Co-creation Data

**IMPORTANT:** This must be done BEFORE scoring to ensure Group and Co-creation columns are available.

```{r group-mapping-and-merge}
# Get participant to group mapping from engagement data
mapping <- df_engagement %>%
  select(`Participant ID`, Group) %>%
  distinct() %>%
  mutate(
    `Participant ID` = as.numeric(`Participant ID`),
    Group = as.numeric(Group)
  ) %>%
  filter(!is.na(`Participant ID`), !is.na(Group))

cat("Group mapping extracted:\n")
cat("  Unique students:", n_distinct(mapping$`Participant ID`), "\n")
cat("  Unique groups:", n_distinct(mapping$Group), "\n\n")

# Clean co-creation column name
cocreation_col <- "Co-creation Level (0:LLM did it all, 1: concepts by students/rest llm, 2: concepts by the students, co-creation of the mnemonic, 3: mainly by students)"

# Prepare co-creation data
df_cocreation_clean <- df_cocreation %>%
  select(Group, Cocreation_Level = all_of(cocreation_col)) %>%
  filter(!is.na(Group), !is.na(Cocreation_Level)) %>%
  mutate(Group = as.numeric(Group))

cat("Co-creation data loaded:\n")
cat("  Groups with co-creation data:", nrow(df_cocreation_clean), "\n\n")
# Merge children data with group mapping, then with co-creation

df_children <- df_children %>%
  mutate(`Participant ID` = as.numeric(`Participant ID`)) %>%
  left_join(mapping, by = "Participant ID") %>%
  filter(!is.na(Group)) %>%
  left_join(df_cocreation_clean, by = "Group") %>%
  filter(!is.na(Cocreation_Level))

cat("Student-level data with co-creation:\n")
cat("  Total observations:", nrow(df_children), "\n")
cat("  Unique students:", n_distinct(df_children$`Participant ID`), "\n")
cat("  Unique groups:", n_distinct(df_children$Group), "\n\n")

# Display group sizes
df_children %>%
  group_by(Group, Timepoint) %>%
  summarise(n_students = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Timepoint, values_from = n_students) %>%
  kable(caption = "Number of Students per Group by Timepoint")
```

## Data Validation and Cleaning

Now check the merged data for any coding issues:

```{r data-validation}
# Check raw values for microcontroller questions
cat("Raw microcontrollers_1 values:\n")
table(df_children$microcontrollers_1, useNA = "ifany")

cat("\nRaw microcontrollers_2 values:\n")
table(df_children$microcontrollers_2, useNA = "ifany")

cat("\nNote: Value 99 = 'I don't know' (no answer attempted)\n")
cat("Value 1 = correct answer\n")
cat("Value 0 = incorrect answer\n")
```

## Create Knowledge Scores

We focus on microcontroller questions as they were the topic of the LLM interaction.

**Scoring rules:**
- Correct answer (value = 1) → 1 point
- Incorrect answer (value = 0) → 0 points  
- "I don't know" (value = 99) → 0 points (treated as incorrect)
- Total score range: 0-2 points

**Note:** We treat "I don't know" (99) as incorrect (0 points) because it indicates lack of knowledge, similar to a wrong answer.

```{r score-knowledge}
# First, let's check for any unexpected values
cat("Checking for unexpected values:\n")
cat("microcontrollers_1 unique values:", 
    unique(as.numeric(df_children$microcontrollers_1)), "\n")
cat("microcontrollers_2 unique values:", 
    unique(as.numeric(df_children$microcontrollers_2)), "\n\n")

# Score microcontroller questions
# Expected values: 1 = correct, 0 = incorrect, 99 = "I don't know"
df_children <- df_children %>%
  mutate(
    # Convert to numeric and check for valid values
    mc1_raw = as.numeric(microcontrollers_1),
    mc2_raw = as.numeric(microcontrollers_2),
    
    # Validate: only 0, 1, or 99 are valid
    mc1_valid = mc1_raw %in% c(0, 1, 99),
    mc2_valid = mc2_raw %in% c(0, 1, 99),
    
    # Score each question: 1 if correct, 0 if incorrect or "don't know"
    mc1_score = if_else(mc1_valid & mc1_raw == 1, 1, 0),
    mc2_score = if_else(mc2_valid & mc2_raw == 1, 1, 0),
    
    # Total score (0-2)
    score = mc1_score + mc2_score,
    
    # Track response types separately for analysis
    n_correct  = if_else(mc1_valid & mc1_raw == 1, 1, 0) + 
                  if_else(mc2_valid & mc2_raw == 1, 1, 0),
    n_dont_know = if_else(mc1_valid & mc1_raw == 99, 1, 0) + 
                  if_else(mc2_valid & mc2_raw == 99, 1, 0),
    n_incorrect = if_else(mc1_valid & mc1_raw == 0, 1, 0) + 
                  if_else(mc2_valid & mc2_raw == 0, 1, 0),
    
    # Flag invalid responses
    has_invalid = !mc1_valid | !mc2_valid
  )

# Report any invalid data
invalid_count <- sum(df_children$has_invalid, na.rm = TRUE)
if (invalid_count > 0) {
  cat("\n[WARNING] Found", invalid_count, "rows with invalid values:\n")
  df_children %>%
    filter(has_invalid) %>%
    select(`Participant ID`, Timepoint, mc1_raw, mc2_raw) %>%
    print()
  cat("\nThese rows will be excluded from analysis.\n\n")
}

# Keep only valid data
df_children_valid <- df_children %>%
  filter(!has_invalid)

# Check what columns we have - DEBUG
cat("\n=== DEBUG: Column check ===\n")
cat("Total columns:", length(names(df_children_valid)), "\n")
cat("Columns in validated data:\n")
print(names(df_children_valid))
cat("\nHas Group column:", "Group" %in% names(df_children_valid), "\n")
cat("Has Participant ID:", "Participant ID" %in% names(df_children_valid), "\n")
cat("Has Timepoint:", "Timepoint" %in% names(df_children_valid), "\n")
cat("Has Cocreation_Level:", "Cocreation_Level" %in% names(df_children_valid), "\n")
cat("========================\n\n")

# Use valid data for all subsequent analyses
df_children <- df_children_valid

# Display response patterns
cat("Response Pattern Summary (Valid Data Only):\n")
cat("==========================================\n\n")

response_summary <- df_children %>%
  group_by(Timepoint) %>%
  summarise(
    n_students = n(),
    mean_score = mean(score, na.rm = TRUE),
    mean_dont_know = mean(n_dont_know, na.rm = TRUE),
    mean_incorrect = mean(n_incorrect, na.rm = TRUE),
    pct_with_dont_know = mean(n_dont_know > 0) * 100, # What % of students had at least one don't know?
    pct_with_incorrect = mean(n_incorrect > 0) * 100, # What % of students had at least one incorrect?
    pct_with_correct = mean(n_correct > 0) * 100,
    pct_dont_know = sum(n_dont_know, na.rm = TRUE) / sum(n_dont_know + n_incorrect + n_correct, na.rm = TRUE) * 100, # What % of all responses were don't know?
    pct_incorrect = sum(n_incorrect, na.rm = TRUE) / sum(n_dont_know + n_incorrect + n_correct, na.rm = TRUE) * 100, # What % of all responses were incorrect?
    pct_correct = sum(n_correct, na.rm = TRUE) / sum(n_dont_know + n_incorrect + n_correct, na.rm = TRUE) * 100,
   )

response_summary %>%
  kable(digits = 2, caption = "Response Patterns by Timepoint")

# Score distribution by timepoint
df_children %>%
  group_by(Timepoint) %>%
  summarise(
    n = n(),
    mean_score = mean(score, na.rm = TRUE),
    sd_score = sd(score, na.rm = TRUE),
    min = min(score, na.rm = TRUE),
    max = max(score, na.rm = TRUE)
  ) %>%
  kable(digits = 2, caption = "Knowledge Scores by Timepoint")

# Detailed breakdown: Number of students per score
cat("\nScore Distribution:\n")
df_children %>%
  group_by(Timepoint, score) %>%
  summarise(n = n(), .groups = 'drop') %>%
  pivot_wider(names_from = score, values_from = n, values_fill = 0) %>%
  kable(caption = "Number of Students by Score and Timepoint")

# Response type breakdown (correct, incorrect, don't know)
cat("\nResponse Type Breakdown (per question):\n")
df_children %>%
  group_by(Timepoint) %>%
  summarise(
    total_questions = n() * 2,
    n_correct = sum(mc1_score + mc2_score),
    n_incorrect = sum(n_incorrect),
    n_dont_know = sum(n_dont_know),
    pct_correct = (n_correct / total_questions) * 100,
    pct_incorrect = (n_incorrect / total_questions) * 100,
    pct_dont_know = (n_dont_know / total_questions) * 100
  ) %>%
  kable(digits = 1, caption = "Percentage of Each Response Type")
```

## Reshape to Long Format (After Validation)

Now that we have clean, validated data, we can reshape for longitudinal analysis:

```{r reshape-long-validated}
# Create long format using VALIDATED data only
# First verify we have all required columns
required_cols <- c("Participant ID", "Group", "Timepoint", "Cocreation_Level", "score")
missing_cols <- setdiff(required_cols, names(df_children))

if (length(missing_cols) > 0) {
  stop("ERROR: Missing required columns: ", paste(missing_cols, collapse = ", "))
}

cat("Creating long format data...\n")
df_long <- df_children %>%
  select(`Participant ID`, Group, Timepoint, Cocreation_Level, score, 
         n_dont_know, n_incorrect, mc1_score, mc2_score) %>%
  mutate(
    # Convert timepoint to numeric (1, 4, 7)
    Time_numeric = case_when(
      Timepoint == "T1" ~ 1,
      Timepoint == "T4" ~ 4,
      Timepoint == "T7" ~ 7,
      TRUE ~ NA_real_
    ),
    # Factor versions for plotting
    Timepoint_factor = factor(Timepoint, levels = c("T1", "T4", "T7")),
    Group_factor = factor(Group),
    Student_ID = factor(`Participant ID`)
  ) %>%
  filter(!is.na(score), !is.na(Time_numeric))

cat("Longitudinal data structure (VALIDATED):\n")
cat("=========================================\n")
cat("Total observations:", nrow(df_long), "\n")
cat("Unique students:", n_distinct(df_long$Student_ID), "\n")
cat("Unique groups:", n_distinct(df_long$Group), "\n")
cat("\nScore range:", min(df_long$score), "to", max(df_long$score), "\n")
cat("Mean score:", round(mean(df_long$score), 2), "\n\n")

summary(df_long %>% select(Time_numeric, Cocreation_Level, score, Group))
```

# Descriptive Statistics

## Co-creation Levels

```{r desc-cocreation}
df_cocreation_clean %>%
  summarise(
    n_groups = n(),
    mean_level = mean(Cocreation_Level),
    sd_level = sd(Cocreation_Level),
    min = min(Cocreation_Level),
    max = max(Cocreation_Level)
  ) %>%
  kable(digits = 2, caption = "Co-creation Level Distribution")

# Distribution plot
ggplot(df_cocreation_clean, aes(x = Cocreation_Level)) +
  geom_histogram(binwidth = 0.5, fill = "#3b82f6", alpha = 0.7) +
  labs(
    title = "Distribution of Co-creation Levels",
    x = "Co-creation Level",
    y = "Number of Groups"
  ) +
  theme_minimal(base_size = 14)

```

## Knowledge Scores Over Time

```{r desc-knowledge}
# Print df_long structure for inspection
cat("\n=== df_long DATA STRUCTURE ===\n")
cat("Dimensions:", nrow(df_long), "rows ×", ncol(df_long), "columns\n\n")
cat("Column names:\n")
print(names(df_long))
cat("\nFirst 20 rows:\n")
print(head(df_long, 20))
cat("\nSummary statistics:\n")
print(summary(df_long))
cat("\n==============================\n\n")

# CHECK FOR OUTLIERS (scores < 0 or > 2)
cat("\n=== CHECKING FOR INVALID SCORES ===\n")
outliers_low <- df_long %>% filter(score < 0)
outliers_high <- df_long %>% filter(score > 2)

if (nrow(outliers_low) > 0) {
  cat("\n[WARNING] Found", nrow(outliers_low), "scores BELOW 0:\n")
  print(outliers_low %>% select(`Participant ID`, Group, Timepoint, score, mc1_raw, mc2_raw, mc1_score, mc2_score))
} else {
  cat("\n[OK] No scores below 0\n")
}

if (nrow(outliers_high) > 0) {
  cat("\n[WARNING] Found", nrow(outliers_high), "scores ABOVE 2:\n")
  print(outliers_high %>% select(`Participant ID`, Group, Timepoint, score, mc1_raw, mc2_raw, mc1_score, mc2_score))
} else {
  cat("\n[OK] No scores above 2\n")
}

cat("\nScore range in df_long:", min(df_long$score), "to", max(df_long$score), "\n")
cat("===================================\n\n")

# Summary by timepoint
df_long %>%
  group_by(Timepoint_factor) %>%
  summarise(
    n_students = n(),
    mean_score = mean(score),
    sd_score = sd(score),
    se_score = sd_score / sqrt(n_students)
  ) %>%
  kable(digits = 2, caption = "Knowledge Scores by Timepoint")

# Plot knowledge evolution
# NOTE: Using geom_jitter with height=0 to avoid y-axis displacement
ggplot(df_long, aes(x = Timepoint_factor, y = score)) +
  geom_boxplot(fill = "#93C5FD", alpha = 0.7) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5, size = 2) +  # height=0 prevents y-axis jitter
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4, 
               fill = "red", color = "red") +
  scale_y_continuous(limits = c(-0.1, 2.1), breaks = seq(0, 2, 0.5)) +  # Set y-axis limits
  labs(
    title = "Knowledge Scores Across Timepoints",
    subtitle = "Red diamond = mean. All scores are between 0-2.",
    x = "Timepoint",
    y = "Knowledge Score (0-2)"
  ) +
  theme_minimal(base_size = 14)
```

## Knowledge by Co-creation Level

```{r desc-cocreation-knowledge}
df_long %>%
  group_by(Cocreation_Level, Timepoint_factor) %>%
  summarise(
    n = n(),
    mean_score = mean(score),
    sd_score = sd(score),
    .groups = 'drop'
  ) %>%
  pivot_wider(
    names_from = Timepoint_factor,
    values_from = c(n, mean_score, sd_score)
  ) %>%
  kable(digits = 2, caption = "Knowledge Scores by Co-creation Level and Timepoint")

# Plot
ggplot(df_long, aes(x = Cocreation_Level, y = score, color = Timepoint_factor)) +
  geom_point(alpha = 0.6, size = 3, position = position_jitter(width = 0.05)) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_color_manual(values = c("T1" = "#3b82f6", "T4" = "#f59e0b", "T7" = "#10b981")) +
  labs(
    title = "Co-creation Level vs Knowledge Score",
    subtitle = "By Timepoint",
    x = "Co-creation Level",
    y = "Knowledge Score",
    color = "Timepoint"
  ) +
  theme_minimal(base_size = 14)
```

# Mixed Effects Models

## Model 1: T7 Only - Co-creation Predicts Learning

Does co-creation level predict post-test knowledge (T7)?

```{r model1-t7}
# This model runs at the student level with group-level random effects.
# Filter T7 data only
df_t7 <- df_long %>% filter(Timepoint == "T7")

# Mixed model with random intercept for groups
model1 <- lmer(
  score ~ Cocreation_Level + (1 | Group_factor),
  data = df_t7,
  REML = TRUE
)

# Summary
summary(model1)

# Tidy output
tidy(model1, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 1: Co-creation → T7 Knowledge")

# Model performance
cat("\nModel Performance:\n")
performance::icc(model1)
performance::r2(model1)
```

### Model 1 Diagnostics

```{r model1-diagnostics, fig.width=12, fig.height=10}
# Check assumptions
check_model(model1, panel = TRUE)

# Additional diagnostic plots
par(mfrow = c(2, 2))
plot(model1, main = "Model 1: Residuals vs Fitted")
qqnorm(resid(model1), main = "Q-Q Plot: Normality of Residuals")
qqline(resid(model1))
hist(resid(model1), main = "Histogram of Residuals", xlab = "Residuals", breaks = 10)
plot(fitted(model1), resid(model1), main = "Fitted vs Residuals", 
     xlab = "Fitted values", ylab = "Residuals")
abline(h = 0, lty = 2, col = "red")
par(mfrow = c(1, 1))
```

### Model 1 Interpretation

```{r model1-interpretation}
cat("\n=== MODEL 1 INTERPRETATION ===\n")
cat("Research Question: Does co-creation level predict T7 knowledge?\n\n")

# Extract key statistics
beta_cocreation <- fixef(model1)["Cocreation_Level"]
se_cocreation <- sqrt(diag(vcov(model1)))["Cocreation_Level"]
t_value <- summary(model1)$coefficients["Cocreation_Level", "t value"]
p_value <- summary(model1)$coefficients["Cocreation_Level", "Pr(>|t|)"]
ci <- confint(model1, parm = "Cocreation_Level", level = 0.95)

cat("Fixed Effects:\n")
cat(sprintf("  Co-creation coefficient (beta): %.3f\n", beta_cocreation))
cat(sprintf("  Standard error: %.3f\n", se_cocreation))
cat(sprintf("  95%% CI: [%.3f, %.3f]\n", ci[1], ci[2]))
cat(sprintf("  t-value: %.3f\n", t_value))
cat(sprintf("  p-value: %.4f\n", p_value))

# Effect size interpretation
cat("\nEffect Interpretation:\n")
if (p_value < 0.05) {
  cat("  [YES] SIGNIFICANT effect of co-creation on T7 knowledge\n")
  cat(sprintf("  - For each 1-unit increase in co-creation level,\n"))
  cat(sprintf("    knowledge score increases by %.2f points (out of 2)\n", beta_cocreation))
  cat(sprintf("  - This represents a %.1f%% increase per co-creation level\n", beta_cocreation/2*100))
} else {
  cat("  [NO] NO significant effect of co-creation on T7 knowledge\n")
}

# Random effects
var_group <- as.data.frame(VarCorr(model1))$vcov[1]
var_resid <- attr(VarCorr(model1), "sc")^2
icc_val <- var_group / (var_group + var_resid)

cat("\nRandom Effects:\n")
cat(sprintf("  Group variance: %.3f\n", var_group))
cat(sprintf("  Residual variance: %.3f\n", var_resid))
cat(sprintf("  ICC (Intraclass Correlation): %.3f\n", icc_val))
cat(sprintf("  - %.1f%% of variance is between groups\n", icc_val*100))
cat(sprintf("  - %.1f%% of variance is within groups (between students)\n", (1-icc_val)*100))

# Model fit
cat("\nModel Fit:\n")
cat(sprintf("  AIC: %.1f\n", AIC(model1)))
cat(sprintf("  BIC: %.1f\n", BIC(model1)))
cat(sprintf("  Log-likelihood: %.2f\n", logLik(model1)))

# R-squared
r2_vals <- r2(model1)
cat(sprintf("  Marginal R² (fixed effects only): %.3f\n", r2_vals$R2_marginal))
cat(sprintf("  Conditional R² (fixed + random): %.3f\n", r2_vals$R2_conditional))

cat("\n================================\n")
```

### Model 1 Visualization

```{r model1-plot, fig.width=10, fig.height=6}
# Plot fixed effects
plot_model(model1, type = "pred", terms = "Cocreation_Level") +
  labs(
    title = "Predicted T7 Knowledge by Co-creation Level",
    subtitle = "Model 1: Mixed effects with random intercepts for groups",
    x = "Co-creation Level",
    y = "Predicted Knowledge Score"
  ) +
  theme_minimal(base_size = 14)
```
### Model 1 : another approach since group random effects at zero
```{r model1_alternative}
# Load required packages
library(sandwich)
library(lmtest)
library(broom)
library(knitr)

# Filter T7 data only
df_t7 <- df_long %>% filter(Timepoint == "T7")

# Fit simple linear regression
model1_simple <- lm(score ~ Cocreation_Level, data = df_t7)

# Get cluster-robust standard errors
model1_robust <- coeftest(model1_simple, vcov = vcovCL, cluster = ~Group_factor)

# Display results
cat("Linear Regression with Cluster-Robust Standard Errors\n")
cat("======================================================\n\n")
print(model1_robust)

# Get confidence intervals with robust SEs
robust_ci <- coefci(model1_simple, vcov = vcovCL, cluster = ~Group_factor)
print(robust_ci)

# Create a tidy table
robust_results <- tidy(model1_robust, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 1: Co-creation → T7 Knowledge (Cluster-Robust SE)")

print(robust_results)

# Model R-squared (not adjusted for clustering, but informative)
cat("\nModel R-squared:", summary(model1_simple)$r.squared, "\n")
cat("Adjusted R-squared:", summary(model1_simple)$adj.r.squared, "\n")
cat("\nNumber of observations:", nobs(model1_simple), "\n")
cat("Number of clusters (groups):", length(unique(df_t7$Group_factor)), "\n")
```
## Model 2: T1 Predicts Co-creation (Selection Effect)

2. Does baseline knowledge (T1) predict co-creation level (selection effect)?


```{r model2-t1}
# Filter T1 data
df_t1 <- df_long %>% filter(Timepoint == "T1")

# Since co-creation is at group level, we need group-level analysis
df_t1_group <- df_t1 %>%
  group_by(Group_factor) %>%
  summarise(
    mean_T1_score = mean(score, na.rm = TRUE),
    Cocreation_Level = first(Cocreation_Level)
  )

# Simple linear model (no random effects needed - one obs per group)
model2 <- lm(Cocreation_Level ~ mean_T1_score, data = df_t1_group)

# Summary
summary(model2)

# Tidy output
tidy(model2, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 2: T1 Knowledge → Co-creation Level")
```

### Model 2 Visualization

```{r model2-plot, fig.width=8, fig.height=6}
ggplot(df_t1_group, aes(x = mean_T1_score, y = Cocreation_Level)) +
  geom_point(size = 4, alpha = 0.7, color = "#3b82f6") +
  geom_smooth(method = "lm", se = TRUE, color = "red", fill = "pink", alpha = 0.3) +
  geom_text(aes(label = Group_factor), nudge_y = 0.1, size = 3) +
  labs(
    title = "Baseline Knowledge vs Co-creation Level",
    subtitle = paste0("beta = ", round(coef(model2)[2], 3), 
                     ", p = ", round(summary(model2)$coefficients[2, 4], 3)),
    x = "Mean T1 Knowledge Score (Baseline)",
    y = "Co-creation Level"
  ) +
  xlim(-0.1, 2) +
  theme_minimal(base_size = 14)
```

### Model 2 Interpretation

```{r model2-interpretation}
cat("\n=== MODEL 2 INTERPRETATION ===\n")
cat("Research Question: Does baseline knowledge predict co-creation level?\n")
cat("(Testing for selection effects)\n\n")

# Extract coefficients
beta_t1 <- coef(model2)["mean_T1_score"]
se_t1 <- summary(model2)$coefficients["mean_T1_score", "Std. Error"]
t_val <- summary(model2)$coefficients["mean_T1_score", "t value"]
p_val <- summary(model2)$coefficients["mean_T1_score", "Pr(>|t|)"]
ci <- confint(model2, "mean_T1_score", level = 0.95)
r_squared <- summary(model2)$r.squared
adj_r_squared <- summary(model2)$adj.r.squared

cat("Regression Results:\n")
cat(sprintf("  Coefficient (beta): %.3f\n", beta_t1))
cat(sprintf("  Standard error: %.3f\n", se_t1))
cat(sprintf("  95%% CI: [%.3f, %.3f]\n", ci[1], ci[2]))
cat(sprintf("  t(%d) = %.3f\n", model2$df.residual, t_val))
cat(sprintf("  p-value: %.4f\n", p_val))
cat(sprintf("  R-squared = %.3f, Adjusted R-squared = %.3f\n", r_squared, adj_r_squared))

cat("\n=== CRITICAL INTERPRETATION ===\n")
if (p_val < 0.05) {
  cat("  [WARNING] SELECTION EFFECT DETECTED\n\n")
  if (beta_t1 > 0) {
    cat("  -> Groups with HIGHER baseline knowledge had HIGHER co-creation\n")
    cat("  -> This suggests selection bias:\n")
    cat("     - Stronger students may have engaged more actively\n")
    cat("     - Co-creation level may reflect prior ability, not just intervention\n")
  } else {
    cat("  -> Groups with LOWER baseline knowledge had HIGHER co-creation\n")
    cat("  -> This suggests compensatory engagement:\n")
    cat("     - Weaker students may have relied more on LLM\n")
  }
  cat("\n  IMPLICATION FOR CAUSAL INFERENCE:\n")
  cat("  The relationship between co-creation and T7 outcomes may be\n")
  cat("  confounded by baseline differences. Consider:\n")
  cat("  - Adding T1 as a covariate in main models\n")
  cat("  - Reporting selection effect in limitations\n")
  cat("  - Using propensity score matching or ANCOVA\n")
} else {
  cat("  [OK] NO SELECTION EFFECT\n\n")
  cat("  -> Baseline knowledge does NOT predict co-creation level\n")
  cat("  -> Co-creation appears independent of prior knowledge\n")
  cat("  -> This supports causal interpretation:\n")
  cat("     - Co-creation -> Learning relationship is not confounded\n")
  cat("     - by baseline group differences\n")
  cat("\n  IMPLICATION FOR CAUSAL INFERENCE:\n")
  cat("  The co-creation effect on T7 likely represents a genuine\n")
  cat("  learning effect rather than selection artifact.\n")
}

cat("\n================================\n")
```

## Model 3: Longitudinal - Time
3. How does knowledge evolve over time (T1 → T4 → T7)?

```{r model3-longitudinal}
# Mixed model with time 
model3 <- lmer(
  score ~ Time_numeric  + (1 | Group_factor) + (1 | Student_ID),
  data = df_long,
  REML = TRUE
)

# Summary
summary(model3)

# Tidy output
tidy(model3, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 3: Knowledge over Time")

# Model performance
cat("\nModel Performance:\n")
performance::icc(model3)
performance::r2(model3)
```

## Model 4: Longitudinal - Time + Co-creation
4. Does the co-creation effect vary by timepoint?

```{r model4-longitudinal with co-creation}
# Mixed model with time and co-creation
model4 <- lmer(
  score ~ Time_numeric + Cocreation_Level + (1 | Group_factor) + (1 | Student_ID),
  data = df_long,
  REML = TRUE
)

# Summary
summary(model4)

# Tidy output
tidy(model4, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 2: Time + Co-creation")

# Model performance
cat("\nModel Performance:\n")
performance::icc(model4)
performance::r2(model4)
```

### Model 4 Diagnostics

```{r model4-diagnostics, fig.width=12, fig.height=10}
check_model(model4, panel = TRUE)

# Additional diagnostics
par(mfrow = c(2, 2))
plot(model4, main = "Model 2: Residuals vs Fitted")
qqnorm(resid(model4), main = "Q-Q Plot")
qqline(resid(model4))
hist(resid(model4), main = "Histogram of Residuals", breaks = 15)
plot(fitted(model4), df_long$score, main = "Observed vs Fitted",
     xlab = "Fitted", ylab = "Observed")
abline(0, 1, col = "red", lty = 2)
par(mfrow = c(1, 1))
```

### Model 4 Interpretation

```{r model4-interpretation}
cat("\n=== MODEL 4 INTERPRETATION ===\n")
cat("Research Question: How do Time and Co-creation affect knowledge longitudinally?\n\n")

# Extract coefficients
beta_time <- fixef(model4)["Time_numeric"]
beta_cocreation <- fixef(model4)["Cocreation_Level"]
ci_time <- confint(model4, parm = "Time_numeric", level = 0.95)
ci_cocreation <- confint(model4, parm = "Cocreation_Level", level = 0.95)
p_time <- summary(model4)$coefficients["Time_numeric", "Pr(>|t|)"]
p_cocreation <- summary(model4)$coefficients["Cocreation_Level", "Pr(>|t|)"]

cat("Fixed Effects:\n")
cat(sprintf("  Time effect (beta): %.3f, 95%% CI [%.3f, %.3f], p = %.4f\n", 
            beta_time, ci_time[1], ci_time[2], p_time))
cat(sprintf("  Co-creation effect (beta): %.3f, 95%% CI [%.3f, %.3f], p = %.4f\n", 
            beta_cocreation, ci_cocreation[1], ci_cocreation[2], p_cocreation))

cat("\nInterpretation:\n")
if (p_time < 0.05) {
  if (beta_time > 0) {
    cat("  [YES] Knowledge INCREASES over time\n")
    cat(sprintf("    - %.3f points increase per timepoint unit\n", beta_time))
  } else {
    cat("  [YES] Knowledge DECREASES over time\n")
  }
} else {
  cat("  [NO] No significant time effect\n")
}

if (p_cocreation < 0.05) {
  cat("  [YES] Co-creation predicts knowledge across all timepoints\n")
  cat(sprintf("    - %.3f points per co-creation level\n", beta_cocreation))
} else {
  cat("  [NO] No significant co-creation effect when accounting for time\n")
}

# Random effects
var_group <- as.data.frame(VarCorr(model4))$vcov[1]
var_student <- as.data.frame(VarCorr(model4))$vcov[2]
var_resid <- attr(VarCorr(model4), "sc")^2

cat("\nRandom Effects (Variance Components):\n")
cat(sprintf("  Group level: %.3f\n", var_group))
cat(sprintf("  Student level: %.3f\n", var_student))
cat(sprintf("  Residual: %.3f\n", var_resid))

total_var <- var_group + var_student + var_resid
cat(sprintf("\nVariance Partitioning:\n"))
cat(sprintf("  Between groups: %.1f%%\n", var_group/total_var*100))
cat(sprintf("  Between students (within groups): %.1f%%\n", var_student/total_var*100))
cat(sprintf("  Residual (within students): %.1f%%\n", var_resid/total_var*100))

# Model fit
r2_vals <- r2(model4)
cat(sprintf("\nModel Fit:\n"))
cat(sprintf("  Marginal R²: %.3f (variance explained by fixed effects)\n", r2_vals$R2_marginal))
cat(sprintf("  Conditional R²: %.3f (variance explained by full model)\n", r2_vals$R2_conditional))
cat(sprintf("  AIC: %.1f, BIC: %.1f\n", AIC(model4), BIC(model4)))

cat("\n================================\n")
```

### Model 4 Visualization

```{r model4-plot, fig.width=12, fig.height=6}
# Plot both effects
p1 <- plot_model(model4, type = "pred", terms = "Time_numeric") +
  labs(title = "Time Effect", x = "Time (Numeric)", y = "Predicted Score") +
  theme_minimal(base_size = 12)

p2 <- plot_model(model4, type = "pred", terms = "Cocreation_Level") +
  labs(title = "Co-creation Effect", x = "Co-creation Level", y = "Predicted Score") +
  theme_minimal(base_size = 12)

grid.arrange(p1, p2, ncol = 2)
```

```{r model4-alternative}
# ============================================
# MODEL 4: LONGITUDINAL ANALYSIS (SIMPLER APPROACH)
# ============================================

library(sandwich)
library(lmtest)
library(broom)
library(knitr)

# Fit simple linear regression with time and co-creation
model4_simple <- lm(score ~ Time_numeric + Cocreation_Level, data = df_long)

# Get cluster-robust standard errors (clustered by Student_ID for repeated measures)
model4_robust <- coeftest(model4_simple, 
                          vcov = vcovCL, 
                          cluster = ~Student_ID)

# Display results
cat("MODEL 4: Longitudinal Analysis - Time + Co-creation Effects\n")
cat("============================================================\n")
cat("Cluster-robust SEs accounting for repeated measures\n\n")
print(model4_robust)

# Get confidence intervals with robust SEs
robust_ci_model4 <- coefci(model4_simple, 
                           vcov = vcovCL, 
                           cluster = ~Student_ID)
cat("\nConfidence Intervals:\n")
print(robust_ci_model4)

# Create a tidy table
robust_results_model4 <- tidy(model4_robust, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 4: Longitudinal Analysis with Cluster-Robust SE")

print(robust_results_model4)

# Model fit statistics
cat("\n=== Model 4 Fit Statistics ===\n")
cat("R-squared:", summary(model4_simple)$r.squared, "\n")
cat("Adjusted R-squared:", summary(model4_simple)$adj.r.squared, "\n")
cat("\nNumber of observations:", nobs(model4_simple), "\n")
cat("Number of students (clusters):", length(unique(df_long$Student_ID)), "\n")
cat("Number of groups:", length(unique(df_long$Group_factor)), "\n")

# Interpretation guide
cat("\n=== Model 4 Interpretation ===\n")
cat("Time_numeric: Change in score per timepoint unit (learning trajectory)\n")
cat("Cocreation_Level: Effect of co-creation level across all timepoints\n")
```

## Model 5: Time × Co-creation Interaction

Does the co-creation effect differ by timepoint?

```{r model5-interaction}
# Mixed model with interaction
model5 <- lmer(
  score ~ Time_numeric * Cocreation_Level + (1 | Group_factor) + (1 | Student_ID),
  data = df_long,
  REML = TRUE
)

# Summary
summary(model5)

# Tidy output
tidy(model5, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 5: Time × Co-creation Interaction")

# Model performance
cat("\nModel Performance:\n")
performance::icc(model5)
performance::r2(model5)

# Test significance of interaction
cat("\nInteraction test:\n")
anova(model4, model5)
```

### Model 5 Visualization

```{r model5-plot, fig.width=10, fig.height=6}
# Plot interaction
plot_model(model5, type = "pred", terms = c("Time_numeric", "Cocreation_Level [0, 1, 2, 3]")) +
  labs(
    title = "Time × Co-creation Interaction",
    subtitle = "Predicted knowledge scores",
    x = "Time (Numeric: 1=T1, 4=T4, 7=T7)",
    y = "Predicted Knowledge Score",
    color = "Co-creation\nLevel"
  ) +
  theme_minimal(base_size = 14)
```

### Model 5 Interpretation

```{r model5-interpretation}
cat("\n=== MODEL 5 INTERPRETATION ===\n")
cat("Research Question: Does the co-creation effect change over time?\n\n")

# Extract coefficients
beta_interaction <- fixef(model5)["Time_numeric:Cocreation_Level"]
ci_interaction <- confint(model5, parm = "Time_numeric:Cocreation_Level", level = 0.95)
p_interaction <- summary(model5)$coefficients["Time_numeric:Cocreation_Level", "Pr(>|t|)"]

cat("Interaction Effect:\n")
cat(sprintf("  Time x Co-creation (beta): %.3f\n", beta_interaction))
cat(sprintf("  95%% CI: [%.3f, %.3f]\n", ci_interaction[1], ci_interaction[2]))
cat(sprintf("  p-value: %.4f\n", p_interaction))

cat("\nInterpretation:\n")
if (p_interaction < 0.05) {
  cat("  [YES] SIGNIFICANT INTERACTION\n")
  if (beta_interaction > 0) {
    cat("  -> The co-creation effect INCREASES over time\n")
    cat("  -> Higher co-creation levels show greater learning gains from T1 to T7\n")
    cat("  -> This suggests co-creation has a cumulative benefit\n")
  } else {
    cat("  -> The co-creation effect DECREASES over time\n")
    cat("  -> Benefits of co-creation diminish at later timepoints\n")
  }
} else {
  cat("  [NO] NO SIGNIFICANT INTERACTION\n")
  cat("  -> The co-creation effect is CONSISTENT across timepoints\n")
  cat("  -> Co-creation has the same impact at T1, T4, and T7\n")
}

# Model comparison
cat("\n Model Comparison:\n")
anova_result <- anova(model4, model5)
p_lr_test <- anova_result$`Pr(>Chisq)`[2]
cat(sprintf("  Likelihood ratio test: chi-squared(1) = %.2f, p = %.4f\n", 
            anova_result$Chisq[2], p_lr_test))

if (p_lr_test < 0.05) {
  cat("  -> Model with interaction fits significantly better\n")
} else {
  cat("  -> Interaction does not significantly improve model fit\n")
  cat("  -> Simpler model (Model 4) is preferred\n")
}

# Effect sizes
r2_vals <- r2(model5)
cat(sprintf("\nModel Fit:\n"))
cat(sprintf("  Marginal R²: %.3f\n", r2_vals$R2_marginal))
cat(sprintf("  Conditional R²: %.3f\n", r2_vals$R2_conditional))

cat("\n================================\n")
```
## Model 6: T4 Predicts Co-creation (Selection Effect)?

Additional question:Does previous knowledge (T4) predict co-creation level (selection effect)?


```{r model6-t4}
# Filter T1 data
df_t4 <- df_long %>% filter(Timepoint == "T4")

# Since co-creation is at group level, we need group-level analysis
df_t4_group <- df_t4 %>%
  group_by(Group_factor) %>%
  summarise(
    mean_T4_score = mean(score, na.rm = TRUE),
    Cocreation_Level = first(Cocreation_Level)
  )

# Simple linear model (no random effects needed - one obs per group)
model6 <- lm(Cocreation_Level ~ mean_T4_score, data = df_t4_group)

# Summary
summary(model6)

# Tidy output
tidy(model6, conf.int = TRUE) %>%
  kable(digits = 3, caption = "Model 2: T1 Knowledge → Co-creation Level")
```


# Model Comparison

```{r model-comparison}
# Compare models
comparison <- compare_performance(model1, model2, model3, model4, model5, verbose = FALSE)

comparison %>%
  kable(digits = 3, caption = "Model Comparison")

# AIC/BIC comparison
cat("\nLower AIC/BIC indicates better model fit\n")
```

# Summary of Results

```{r summary-table}
# Create summary table
results_summary <- tibble(
  Model = c(
    "Model 1: Co-creation predicting T7",
    "Model 2: T1 predicting Co-creation",
    "Model 3: Longitudinal",
    "Model 4: Longitudinal with co-creation",
    "Model 5: Interaction"
  ),
  Formula = c(
    "Score ~ Co-creation + (1|Group)",
    "Co-creation ~ T1_Score",
    "Score ~ Time + (1|Group) + (1|Student)",
    "Score ~ Time + Co-creation + (1|Group) + (1|Student)",
    "Score ~ Time × Co-creation + (1|Group) + (1|Student)"
  ),
  
  Key_Finding = c(
  # Model 1: mixed model (T7 only)
  paste0("Co-creation beta = ", round(fixef(model1)[2], 3), 
         ", p = ", format.pval(coef(summary(model1))[2, 5], digits = 3)),
  # Model 2: simple lm (T1 -> Co-creation)
  paste0("T1 beta = ", round(coef(summary(model2))[2, 1], 3), 
         ", p = ", format.pval(coef(summary(model2))[2, 4], digits = 3)),
  # Model 3: mixed model (time only)
  paste0("Time beta = ", round(fixef(model3)[2], 3), 
         ", p = ", format.pval(coef(summary(model3))[2, 5], digits = 3)),
  # Model 4: mixed model (time + co-creation)
  paste0("Co-creation beta = ", round(fixef(model4)[3], 3), 
         ", p = ", format.pval(coef(summary(model4))[3, 5], digits = 3)),
  # Model 5: mixed model (interaction)
  paste0("Interaction beta = ", round(fixef(model5)[4], 3), 
         ", p = ", format.pval(coef(summary(model5))[4, 5], digits = 3))
)
)

results_summary %>%
  kable(caption = "Summary of All Models")

```

# Interpretation

## Key Findings

```{r interpretation}
# Extract key statistics
t7_effect <- fixef(model1)["Cocreation_Level"]
t7_p <- summary(model1)$coefficients["Cocreation_Level", "Pr(>|t|)"]

long_effect <- fixef(model4)["Cocreation_Level"]
long_p <- summary(model4)$coefficients["Cocreation_Level", "Pr(>|t|)"]

t1_effect <- coef(model2)["mean_T1_score"]
t1_p <- summary(model2)$coefficients["mean_T1_score", "Pr(>|t|)"]

cat("=== INTERPRETATION ===\n\n")

cat("1. CO-CREATION EFFECT ON LEARNING (T7):\n")
cat(sprintf("   beta = %.3f, p = %.4f\n", t7_effect, t7_p))
if (t7_p < 0.05) {
  cat("   [YES] SIGNIFICANT: Higher co-creation predicts better post-test knowledge\n")
} else {
  cat("   [NO] NOT SIGNIFICANT\n")
}

cat("\n2. LONGITUDINAL CO-CREATION EFFECT:\n")
cat(sprintf("   beta = %.3f, p = %.4f\n", long_effect, long_p))
if (long_p < 0.05) {
  cat("   [YES] SIGNIFICANT: Co-creation effect holds across all timepoints\n")
} else {
  cat("   [NO] NOT SIGNIFICANT: Effect may be specific to T7\n")
}

cat("\n3. SELECTION EFFECT (T1 -> Co-creation):\n")
cat(sprintf("   beta = %.3f, p = %.4f\n", t1_effect, t1_p))
if (t1_p < 0.05) {
  cat("   [WARNING] SIGNIFICANT: Baseline knowledge predicts co-creation level\n")
  cat("   This suggests potential selection effects\n")
} else {
  cat("   [OK] NOT SIGNIFICANT: No evidence of selection bias\n")
  cat("   Co-creation appears independent of baseline knowledge\n")
}
```

## Recommended Reporting

```{r reporting}
cat("\n=== SUGGESTED TEXT FOR PAPER ===\n\n")

cat("We examined the relationship between co-creation level during LLM interaction\n")
cat("and learning outcomes using mixed-effects models to account for students nested\n")
cat("within groups.\n\n")

if (t7_p < 0.05) {
  cat(sprintf("Co-creation level significantly predicted post-test knowledge (T7)\n"))
  cat(sprintf("(beta = %.2f, 95%% CI [%.2f, %.2f], p = %.3f).\n", 
              t7_effect, 
              confint(model1)["Cocreation_Level", 1],
              confint(model1)["Cocreation_Level", 2],
              t7_p))
}

cat("\nThe mixed model showed a singular fit with zero group-level variance,\n")
cat("indicating that variability in knowledge scores is primarily within-group\n")
cat("rather than between-group differences.\n\n")

if (t1_p >= 0.05) {
  cat("\nImportantly, baseline knowledge (T1) did not predict co-creation level\n")
  cat(sprintf("(beta = %.2f, p = %.3f), suggesting that the observed co-creation effect\n", 
              t1_effect, t1_p))
  cat("represents a genuine learning effect rather than a selection artifact.\n")
}
```

